### 2020-12-22 papers/spherical-caps/highlights.tex
Add statement on significance and highlights

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex

\graphicspath{ {images/} }

\include{bsthesisnewcommands}
\include{bsmacros}
\include{somacros}

\title{Sparse spectral methods for partial differential equations on spherical caps, highlights}
\author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\begin{enumerate}
\item Constructs sparse discretisations for partial differential operators on spherical caps using orthogonal polynomials. 
\item Approximations are observed to converge spectrally fast.
\item Optimal complexity is achieved for operators that are invariant to rotation around the pole.
\item Examples include (surface) Poisson equation, Helmholtz equation with variable coefficients, and the biharmonic equation.
\end{enumerate}

\end{document}