### 2021-02-17 papers/thesis/full/acknowledgements/acknowledgements.tex

\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

Firstly, I would like to express my thanks to ESPRC for their financial support, and to the Mathematics of Planet Earth Centre for Doctoral Training (MPE CDT) for providing me with the incredible opportunities that the course has brought.

I would like to thank the MPE CDT staff for any knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for his guidance, understanding and general enthusiasm (it meant a lot). And to all my fellow 'Charlies' -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.

I thank my family and friends for all their support and words of encouragement -- it all helped a lot. An extra thanks to all my undergrad mates for many well needed fun times too.

A special thanks goes to my partner Jessica for putting up with me and helping me when things got tough during my 4 years at Imperial College London.

Finally, to Dr. Sheehan Olver -- I cannot thank you enough for all your help, wisdom, advice, kindness, understanding and patience you have given me. I may not have expressed my gratitude enough over the past few years, but you have truly been a great supervisor to me.

I would like to express (whatever feelings I have) to:

\begin{itemize}
 \item My supervisor
 \vspace*{3mm}
 \item My second supervisor
 \vspace*{3mm}
 \item Other researchers
 \vspace*{3mm}
 \item My family and friends
\end{itemize}

\end{acknowledgements}