### 2021-02-17 papers/thesis/full/declaration/declaration.tex

\addcontentsline{toc}{chapter}{Declaration}

\begin{declaration}
	Declaration here.
\end{declaration}