### 2021-02-17 papers/thesis/full/foreword/foreword.tex

\addcontentsline{toc}{chapter}{Foreword}

\begin{foreword}

Text of the foreword.

\end{foreword}