### 2021-04-15 papers/spherical-caps/submitted/significance.tex
Moved submitted SC paper into new directory. Revised version to be added.

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex

\graphicspath{ {images/} }

\include{bsthesisnewcommands}
\include{bsmacros}
\include{somacros}

\title{Sparse spectral methods for partial differential equations on spherical caps, significant and novelty}
\author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

This paper introduces a novel approach for solving partial differential equations on an important subset of the full sphere: spherical caps. The approach leads to sparse discretisations of general partial differential operators that can be efficiently solved with spectral convergence, and in the case of operators that are invariant to rotation around the pole the resulting linear systems can be constructed and solved in optimal complexity. There is a big desire for more localised approaches to solving PDEs on subsections in weather prediction and other applications, and this manuscript lays the groundwork for high order, high accuracy new numerical methods.

\end{document}