### 2021-09-28 papers/thesis/resubmitted/foreword/foreword.tex
Resubmitted version of thesis to track the changes for the corrections. Fixed typo in SH clenshaws.

\cleardoublepage
\addcontentsline{toc}{chapter}{Foreword}

\begin{foreword}

Much of the work in this thesis has already been published, or submitted for publication.

\begin{itemize}

\item The contents of \bsrefchapter{CHAPTER:diskslice} has been published in:

Ben Snowball and Sheehan Olver. Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums. \textit{Studies in Applied Mathematics}, 145(1):3--35, 2020.

DOI: https://doi.org/10.1111/sapm.12303

\item The contents of \bsrefchapter{CHAPTER:sphericalcaps} has been submitted for publication in \textit{Transactions of Mathematics and Its Applications}:

Ben Snowball and Sheehan Olver. Sparse spectral methods for partial differential equations on spherical caps. arXiv preprint arXiv:2012.11493, 2020.

\end{itemize}

\end{foreword}