### 2021-11-09 papers/thesis/resubmitted/acknowledgements/acknowledgements.tex
Actually, spotted typo in EPSRC. Fixed.

\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

Firstly, I would like to express my thanks to EPSRC for their financial support, and to the Mathematics of Planet Earth Centre for Doctoral Training (MPE CDT) for providing me with the incredible opportunities that the course has brought.

I would like to thank the MPE CDT staff for the knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for your guidance, assistance and general enthusiasm. And to all my fellow \enquote{Charlies} -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.

I thank my friends for their support and the fun times they have provided me with over these few years. I also thank my family for all their support and encouragement, not just during my PhD but throughout my life as well to help me get to this point.

A special thanks goes to my wonderful partner Jess for putting up with me and helping me when things got tough during my time at Imperial College London.

Finally, to Dr. Sheehan Olver -- I cannot thank you enough for all your help, wisdom, advice, kindness, understanding and patience you have given me. I may not have expressed my gratitude enough, but you have truly been a great supervisor to me.

%I would like to express (whatever feelings I have) to:
%
%\begin{itemize}
% \item My supervisor
% \vspace*{3mm}
% \item My second supervisor
% \vspace*{3mm}
% \item Other researchers
% \vspace*{3mm}
% \item My family and friends
%\end{itemize}

\end{acknowledgements}